---
title: "Ideology"
author: "Mark Hand"
date: "1/22/2019"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message=FALSE, warning=FALSE)
```

```{r packages}
library(tidyverse) # for data wrangling and plotting 
library(lubridate) # for dates
library(ggjoy) # for kicks (almost) 
```

```{r data}
# Loading DIME data 
# Source http://web.stanford.edu/~bonica/data.html
# Blanked out to enable knitting 
# PoliticalIdeology <- read_csv("data/dime_cong_elections_current.csv")
PoliticalIdeology %>% transform(total_receipts = as.numeric(total_receipts)) -> PoliticalIdeology
PoliticalIdeology %>% filter(cycle==2014) -> PoliticalIdeology2014

# A histogram of Political Ideology in 2014, for both house and senate candidates 
ggplot(subset(PoliticalIdeology2014, seat = federal:house)) +
  geom_histogram(aes(recipient_cfscore)) +
  theme_minimal()

ggplot(subset(PoliticalIdeology2014, seat = federal:senate)) +
  geom_histogram(aes(recipient_cfscore)) +
  theme_minimal()

# Examining 2014 Contribution data
# bigfile.sample <- read.csv("Political_Data/contribDB_2014.csv", # stringsAsFactors=FALSE, header=T, nrows=20)

# Loading 2014 Contribution data 
# Commented out to allow knitting 
# contribDB_2014 <- fread("data/contribDB_2014.csv", select =  c("bonica.rid","transaction.id", "amount","date","contributor.name","contributor.zipcode","recipient.name","seat","contributor.cfscore", "candidate.cfscore", "contributor.state", "recipient.state"))
# head(contribDB_2014)

# Creating Time Variables in Contributions 
contribDB_2014 %>% 
  mutate(date = as.Date(date),
    month = month(date),
    week = week(date),
    year = year(date)) -> contribDB_2014

## Sample Candidate Filters
# table(PoliticalIdeology2014$recipient.name)
# Candidate.Funding_Alquist <- filter(contribDB_2014, recipient.name == "alquist, elaine")
# Candidate.Funding_Beto <- filter(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")

## Looking at data 
# table(contribDB_2014$recipient.state)
# ... strange results. What are these numbers? 
# table(contribDB_2014$contributor.state)
# ... shows lots of types of TX.
# contribDB_2014 %>%
#   mutate(contributor.state = str_replace(contributor.state, c("tx","Tx"), "TX")) -> contribDB_2014_test
# table(contribDB_2014_test$contributor.state)
# Didn't work. What the devil?
# str(contribDB_2014_test)
# contribDB_2014 %>% mutate(contributor.state = recode(contributor.state, "Tx"= "TX",  "tx"="TX")) -> contribDB_2014_test
# table(contribDB_2014_test$contributor.state)
# Worked. So... 
# rm(contribDB_2014_test)

# Tidying TX labels 
contribDB_2014 %>% mutate(contributor.state = recode(contributor.state, "Tx"= "TX",  "tx"="TX")) -> contribDB_2014

#Create TX contributors database 
contribDB_2014_TX <- filter(contribDB_2014, recipient.state == "TX" | contributor.state == "TX")
# ... where "|" means "or"

# Create Beto database 
contribDB_2014_Beto <- filter(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")

# Create TX weekly mean and SD   
contribDB_2014_TX %>% 
  group_by(bonica.rid, recipient.name,seat,candidate.cfscore, year, month, week) %>% 
  summarise(
    contributor.mean.week = mean(contributor.cfscore, na.rm=TRUE),
    contributor.sd.week = sd(contributor.cfscore, na.rm=TRUE)
    ) -> TX2014_weekly

# Create TX monthly mean and SD   
contribDB_2014_TX %>% 
  group_by(bonica.rid, recipient.name,seat,candidate.cfscore, year, month) %>% 
  summarise(
    contributor.mean.month = mean(contributor.cfscore, na.rm=TRUE),
    contributor.sd.month = sd(contributor.cfscore, na.rm=TRUE)
    ) -> TX2014_monthly
```

```{r Beto plots}
# Beto2014 
ggplot(contribDB_2014_Beto) + 
  geom_point(mapping = aes(x = date, y = contributor.cfscore
  # ,size = amount
  )) + theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))

# Beto 2014 by week?  
ggplot(contribDB_2014_Beto) + 
  geom_point(aes(x = week, y = contributor.cfscore
  # ,size = amount
  )) + theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))

# Beto 2014 boxplot by month
ggplot(contribDB_2014_Beto) + 
  geom_boxplot(aes(x = month, y = contributor.cfscore, group = month)) + 
  theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))

# Beto 2014 boxplot by week 
ggplot(contribDB_2014_Beto) + 
  geom_boxplot(aes(x = week, y = contributor.cfscore, group = week)) + 
  theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))

# Beto violin by month 
ggplot(contribDB_2014_Beto) + 
  geom_violin(aes(x = month, y = contributor.cfscore, group = month)) + 
  theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))

# Beto violin by month, weighted by amount 
ggplot(contribDB_2014_Beto) + 
  geom_violin(aes(x = month, y = contributor.cfscore*amount, group = month)) + 
  theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))


# Beto violin by week 
ggplot(contribDB_2014_Beto) + 
  geom_violin(aes(x = week, y = contributor.cfscore, group = week)) + 
  theme_minimal() + labs(list(title='', x='time', y='contributor.cfscore'))

# Beto2014 trend line
ggplot(contribDB_2014_Beto) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal()

# Betojoy fail 
# https://simplystatistics.org/2017/07/13/the-joy-of-no-more-violin-plots/
# ggplot(contribDB_2014_Beto, aes(x = contributor.cfscore, y = month, group = month)) + geom_joy()
```

```{r examining means and sds of TX contributors ideology}
# Looks like I need to filter for federal House and Senate--I had no idea the others were in here! 
TX2014_weekly %>%  filter(seat == "federal:house" | seat == "federal:senate") -> TX2014_weekly

TX2014_monthly %>%  filter(seat == "federal:house" | seat == "federal:senate") -> TX2014_monthly

# A histogram of donor ideologies (mean by month)
ggplot(TX2014_monthly) +
  geom_histogram(aes(contributor.mean.month)) +
  theme_minimal()

# A histogram of donor ideologies (sd by month)
ggplot(TX2014_monthly) +
  geom_histogram(aes(contributor.sd.month)) +
  theme_minimal()

# A histogram of donor ideologies (mean by week)
ggplot(TX2014_weekly) +
  geom_histogram(aes(contributor.mean.week)) +
  theme_minimal()

# A histogram of donor ideologies (sd by week)
ggplot(TX2014_weekly) +
  geom_histogram(aes(contributor.sd.week)) +
  theme_minimal()


```

Ok, so what I think I've done here: I've shown that I can summarize mean and standard deviation of donor ideologies for one candidate, and then for all candidates (in Texas). If that's right, then we should be able to look at what the ideology of early donors (in mean and in standard deviation) is predictive of. 

Asks Peter: "Can you check a simple correlation run a regression without controls on candidate variance and DV of outcome?"

Quick answer: Not with this data, but could perhaps do so with Congressional candidates from multiple years, or maybe state level candidates over multiple years. 

