---
title: "Political Effectuation - Sequence and Stability"
author: "Peter Boumgarden and Mark Hand"
date: "5/24/2018"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message=FALSE, warning=FALSE)
```

## Introduction

The following document walks through an analysis of political effectuation in the form of ideological exploration for candidates for senate and house in the US political races, in particular exploring whether they funding maintained stability or had any sort of sequence in ideological homogeneity.

```{r packages comment=NA, warning=FALSE, message=FALSE} 
# Installing Necessary Packages
# source("http://washuresearch.org/stockfunctions.R") 
library(tidyverse) 
library(ff)
library(data.table)
library(lme4)
```

## Data Cleaning 
```{r comment=NA, warning=FALSE, message=FALSE} 
# Sources 
# https://data.stanford.edu/dime#download-data
# https://elections.huffingtonpost.com/pollster/2014-national-house-race

# What's this? 
# fread("Political_Data/dime_cong_elections_current.csv", select = c(1:3))

#Political ideology 
PoliticalIdeology <- read_csv("data/dime_cong_elections_current.csv")
PoliticalIdeology %>% transform(total_receipts = as.numeric(total_receipts)) -> PoliticalIdeology
PoliticalIdeology %>% filter(cycle==2014) -> PoliticalIdeology2014 

# Examining small number of rows in a large file: 
# bigfile.sample <- read.csv("Political_Data/contribDB_2014.csv", # stringsAsFactors=FALSE, header=T, nrows=20)

# Loading 2014 contributions 
contribDB_2014 <- fread("data/contribDB_2014.csv", select =  c("bonica.rid","transaction.id", "amount","date","contributor.name","contributor.zipcode","recipient.name","seat","contributor.cfscore", "candidate.cfscore", "contributor.state", "recipient.state"))
head(contribDB_2014)

# Setting months and years in contributions 
contribDB_2014 %>% 
  mutate(date = as.Date(date),
    month = month(date),
    year = year(date)) -> contribDB_2014

# Sample Candidate Filters
# Candidate.Funding_Alquist <- filter(contribDB_2014, recipient.name == "alquist, elaine")
# Candidate.Funding_Beto <- filter(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")

# Sample State Filters 
unique(contribDB_2014$recipient.state)
table(contribDB_2014$contributor.state)
contribDB_2014_TX <- filter(contribDB_2014, recipient.state == "TX" | contributor.state == "TX") 

```

## Charts 
```{r some fun charts, include=F}
# http://r4ds.had.co.nz/data-visualisation.html

# All data, by party
ggplot(data = PoliticalIdeology) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=party))

# All data, by win/loss
ggplot(data = PoliticalIdeology) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=gwinner))

# Republican winners and losers 
ggplot(data = subset(PoliticalIdeology, party=="R")) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=gwinner))

# Subset
ggplot(data = subset(PoliticalIdeology, total_receipts > 400000)) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=gwinner))

# Texas 2018 
ggplot(data = subset(PoliticalIdeology, state=="TX" & cycle=="2018")) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=party))

# Histo 
ggplot(data = PoliticalIdeology) +
  geom_histogram(mapping = aes(x = recipient_cfscore))

# Beto2014 
ggplot(data = subset(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")) + 
  geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + theme_minimal() + labs(list(title='', x='', y='')) + theme(axis.ticks = element_blank(), axis.text.x = element_blank()) 

# Beto2014 trend line
ggplot(data = subset(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

# Beto2014 abline
ggplot(data = subset(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal()

# grep('SCALISE', contribDB_2014$recipient.name) 
contribDB_2014$recipient.name[660490]

"SCALISE, STEPHEN J"
ggplot(data = subset(contribDB_2014, recipient.name == "SCALISE, STEPHEN J")) + 
  geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + theme_minimal() + labs(list(title='', x='', y='')) + theme(axis.ticks = element_blank(), axis.text.x = element_blank()) 

ggplot(data = subset(contribDB_2014, recipient.name == "SCALISE, STEPHEN J")) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

ggplot(data = subset(contribDB_2014, recipient.name == "PELOSI, NANCY")) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 
```

### Generic Ballot History 
```{r generic ballot}
generic_ballot_history <- read_csv("data/generic-ballot-history.csv")
View(generic_ballot_history)

ggplot(data = generic_ballot_history) + 
  geom_smooth(mapping = aes(x = end_date, y = Democrat), color="blue") + theme_minimal() + labs(list(title='', x='', y='')) + theme(axis.ticks = element_blank(), axis.text.x = element_blank())

ggplot(data = generic_ballot_history) + theme_minimal() + 
  geom_point(mapping = aes(x = end_date, y = Democrat), color="blue") +   
  geom_smooth(mapping = aes(x = end_date, y = Democrat), color="blue") +   
  geom_point(mapping = aes(x = end_date, y = Republican), color="red") +     
  geom_smooth(mapping = aes(x = end_date, y = Republican), color="red") + 
  geom_point(mapping = aes(x = end_date, y = `Undecided/Other`), color="yellow") + 
  geom_smooth(mapping = aes(x = end_date, y = `Undecided/Other`), color="yellow") + 
  labs(list(title='Generic Polling, 2012-2018', x='', y='% Support'))

ggplot(data = generic_ballot_history) + theme_minimal() + 
  geom_smooth(mapping = aes(x = end_date, y = Democrat), color="blue") +   
  geom_smooth(mapping = aes(x = end_date, y = Republican), color="red") + 
  geom_smooth(mapping = aes(x = end_date, y = `Undecided/Other`), color="yellow") + 
  labs(list(title='Generic Polling, 2012-2018', x='', y='% Support'))
```

### Starting Hypothesis 1: Consistency -> Victory 
Columns to build: 
* Campaign ID (observation)
* Won primary? 0/1 (DV) 
* Won general? 0/1 (DV) <-- might want to separate out, or determine hierarchy 
* Slope of of ideological trend line (how to build?) 
* Amount raised (control)
* Party (control) 
* Tighteness of race (how?)
* Party (control)
* State (fixed effects?)
* Challenger v incumbent
* House v Senate 
* Other controls, per Bonica? 

## Basic modeling 
```{r basic modeling}
# Base model for 2014 cycle 
table(PoliticalIdeology2014$pwinner) 
# Replacing Ls and Ws with 1s and 0s, making things into the right data types   
PoliticalIdeology2014 %>% 
  mutate(pwinner = stringr::str_replace(pwinner, "L", "0"),
         pwinner = stringr::str_replace(pwinner, "W", "1"),
         pwinner = as.factor(pwinner),
         gwinner = as.factor(gwinner),
         ppct = as.numeric(ppct),
         gpct = as.numeric(gpct)) -> PoliticalIdeology2014

# Starting out with disbursements 
summary(glm(pwinner~total_disbursements, data = PoliticalIdeology2014, family = binomial))
summary(glm(gwinner~total_disbursements, data = PoliticalIdeology2014, family = binomial))

# Plots of money raised versus winning primary
ggplot(data=subset(PoliticalIdeology2014, ppct>0.0)) + theme_minimal() + geom_point(mapping = aes(x = total_disbursements, y = ppct, color = pwinner))

# Plots of money raised versus winning general 
ggplot(data=subset(PoliticalIdeology2014, gpct>0.3 & gpct <0.7)) + theme_minimal() + geom_point(mapping = aes(x = total_disbursements, y = gpct, color = gwinner)) + scale_y_continuous()

summary(glm(pwinner~total_disbursements+Incum_Chall+party+recipient_cfscore*party, data = PoliticalIdeology2014, family = binomial))
# More money is good. 
# Being an incumbent is good. 
# Being a Republican in this cycle was bad. 
# For Democrats, being more conservative (having a higher recipient_cfscore) was bad. 
# Gender needs to be (re)constructed. 
# State fixed effects? 

# Filtering for house only 
# PoliticalIdeology2014 %>% filter(seat=="federal:house") -> PoliticalIdeology2014H
# summary(glm(pwinner~total_disbursements+Incum_Chall+party+recipient_cfscore*party, data = PoliticalIdeology2014H, family = binomial))
# Basically the same results 
```

### Trying to get slope of lines
```{r slopes ignore the comments}

# H: Candidates that show greater ideological consistency are more likely to win. 
# DV: Winning
# IV: Slope of ideological line over time 
# Controls in data we have: 
# R v D candidate, generic R v D ballot, state, year, amount raised 
# Controls we need to find: 
# Challenger v. incumbent (though we could look at previous years, closeness of race 

# Works! Now, how to run this over every candidate and then save to a dataframe? This is either a "for "for loop" or create-your-own-function situation. 
# Maybe helpful: https://github.com/tidyverse/dplyr/issues/2177

# For loop fail 
# output <- vector("double", ncol(df)) 
# for (i in seq_along(unique(contribDB_2014_TX$bonica.rid))) 
#  {output[[i]] <- coef(lm(date ~ contributor.cfscore, data = contribDB_2014_TX, subset=(bonica.rid == [[i]])))
#  }

# Function fails  
# x <- unique(contribDB_2014_TX$bonica.rid)
# coef(lm(date ~ contributor.cfscore, data = contribDB_2014_TX, subset=(bonica.rid == x)))

# Function almost works? 
# coef01 <- function (x) { 
#   number_entries <- nrow(contribDB_2014_TX[contribDB_2014_TX$recipient.name ==x, ])
#   print(x)
#   print(number_entries)
#   if (length(!is.na(contribDB_2014_TX$date[contribDB_2014_TX$recipient.name ==x])) > 3 & length(!is.na(contribDB_2014_TX$contributor.cfscore[contribDB_2014_TX$recipient.name ==x])) > 3) {
#     coef(lm(date ~ contributor.cfscore, data = contribDB_2014_TX, subset=(recipient.name == x)))
#   }
# }

# coef01 <- function (x) { 
#   print(x)
#   if (abs(sum(contribDB_2014_TX$contributor.cfscore[contribDB_2014_TX$recipient.name ==x])) > 0) {
#     coef(lm(date ~ contributor.cfscore, data = contribDB_2014_TX, subset=(recipient.name == x)))
#   }
# }

# coef01(x = "FLUOR CORPORATION POLITICAL ACTION COMMITTEE (FLUOR PAC)")
# coef01(x = "O ROURKE, ROBERT (BETO)")
# coef01(x = "TEAMSTERS LOCAL 745 DRIVE")

# contribDB_2014_TX[contribDB_2014_TX$recipient.name =="TEAMSTERS LOCAL 745 DRIVE",]
# length(contribDB_2014_TX$contributor.cfscore[contribDB_2014_TX$recipient.name =="O ROURKE, ROBERT (BETO)"])

# length(!is.na(contribDB_2014_TX$date[contribDB_2014_TX$recipient.name =="TEAMSTERS LOCAL 745 DRIVE"]))
# length(!is.na(contribDB_2014_TX$contributor.cfscore[contribDB_2014_TX$recipient.name =="TEAMSTERS LOCAL 745 DRIVE"]))

# Basic function, for Beto 
contribDB_2014_TX %>% 
  filter(recipient.name == "O ROURKE, ROBERT (BETO)")
coef(lm(date ~ contributor.cfscore, data = contribDB_2014_TX, subset=(bonica.rid == "cand44995")))
# So this is different from the recipient.name function. Why? 

# Function - success! 
coef02 <- function (x) { 
  coef(lm(date ~ contributor.cfscore, data = contribDB_2014_TX02, subset=(bonica.rid == x)))}

# Filtering out NAs becuase they break the function 
contribDB_2014_TX %>% 
  filter(!is.na(contributor.cfscore)) -> contribDB_2014_TX02

# Building list of uniques to run the function on 
TX_uniques02 <- unique(contribDB_2014_TX02$bonica.rid)

# Testing the series 
newdataframe <- data.frame(bonica.rid = TX_uniques02[1:5])
head(newdataframe)
newdataframe$intercept <- sapply(TX_uniques02[1:5], function (i) coef02(i)[1])
newdataframe$slope <- sapply(TX_uniques02[1:5], function (i) coef02(i)[2])
head(newdataframe)

# Running the above on the larger frame 
newdataframe <- data.frame(bonica.rid = TX_uniques02)
head(newdataframe)
newdataframe$intercept <- sapply(TX_uniques02, function (i) coef02(i)[1])
newdataframe$slope <- sapply(TX_uniques02, function (i) coef02(i)[2])
head(newdataframe)

politislope <- left_join(newdataframe, PoliticalIdeology2014, by = c("bonica.rid" = "bonica_rid"))
politislope %>% 
  filter(!is.na(cycle)) -> candislope
# How did we end up with only 99 candidates? Something went wrong, probably with the NAs. 

summary(glm(pwinner~slope*party, data = candislope, family = binomial))
summary(glm(gwinner~slope*party, data = candislope, family = binomial))
summary(glm(pwinner~slope+party, data = candislope, family = binomial))
summary(glm(gwinner~slope+party, data = candislope, family = binomial))
# No results 

```

# Trying out LMER
...  this is too large a file for me to work from. 
```{r lme4}
joined.2014 <- left_join(contribDB_2014, PoliticalIdeology2014, by = c("bonica.rid" = "bonica_rid"))

fm1 <- lmer(pwinner ~ date + (date | contributor.cfscore), joined.2014)
```

### Bibliography test 
[@marsh_catabolic_1981]