 ---
title: "Political Effectuation - Sequence and Stability"
author: "Peter Boumgarden and Mark Hand"
date: "5/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message=FALSE, warning=FALSE)
```

## Introduction

The following document walks through an analysis of political campaign strategy in the form of ideological exploration for candidates for Senate and House political races in the US; in particular, we explore whether the ideology of their funding maintained stability or had any sort of sequence in ideological 'spread.'

```{r comment=NA, warning=FALSE, message=FALSE} 
# Installing Necessary Packages
source("http://washuresearch.org/stockfunctions.R") 
library(tidyverse) 
library(data.table)
library(gtable)
```

## Data Cleaning 
```{r comment=NA, warning=FALSE, message=FALSE} 
# Sources 
# https://data.stanford.edu/dime#download-data
# https://elections.huffingtonpost.com/pollster/2014-national-house-race

# What's this? An approach to big data
# fread("Political_Data/dime_cong_elections_current.csv", select = c(1:3))

#Political ideology 
# PoliticalIdeology <- read_csv("Political_Data/dime_cong_elections_current.csv")
PoliticalIdeology %>% transform(total_receipts = as.numeric(total_receipts)) -> PoliticalIdeology

# Examining small number of rows in a large file: 
# bigfile.sample <- read.csv("Political_Data/contribDB_2014.csv", stringsAsFactors=FALSE, header=T, nrows=20)

# Loading 2014 contributions 
# contribDB_2014 <- fread("Political_Data/contribDB_2014.csv", select =  c("transaction.id", "amount","date","contributor.name","contributor.zipcode","recipient.name","seat","contributor.cfscore", "candidate.cfscore", "contributor.state", "recipient.state"))
# head(contribDB_2014)

# Loading Candidate - Ideology File
DIMEscores <- read_csv("Political_Data/dime_recipients_all_1979_2014.csv")

# Setting months and years in contributions 
contribDB_2014 %>% 
  mutate(date = as.Date(date),
    month = month(date),
    year = year(date)) -> contribDB_2014

# Create a new variable which identifies whether funding comes from within the state

contribDB_2014 %>% 
 mutate(SSfunding = ifelse("contributor.state == recipient.state"),1,0) -> contribDB_2014


# Sample Candidate Filters
Candidate.Funding_Alquist <- filter(contribDB_2014, recipient.name == "alquist, elaine")
Candidate.Funding_Beto <- filter(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")

# Sample State Filters 
unique(contribDB_2014$recipient.state)
contribDB_2014_TX <- filter(contribDB_2014, recipient.state == "TX") 

```

## Charts 
```{r some fun charts, include=F}
# http://r4ds.had.co.nz/data-visualisation.html

# All data, by party
ggplot(data = PoliticalIdeology) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=party))

# All data, by win/loss
ggplot(data = PoliticalIdeology) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=gwinner))

# Republican winners and losers 
ggplot(data = subset(PoliticalIdeology, party=="R")) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=gwinner))

# Subset
ggplot(data = subset(PoliticalIdeology, total_receipts > 400000)) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=gwinner))

# Texas 2018 
ggplot(data = subset(PoliticalIdeology, state=="TX" & cycle=="2018")) + 
  geom_point(mapping = aes(x = recipient_cfscore , y = total_receipts, color=party))

# Histo 
ggplot(data = PoliticalIdeology) +
  geom_histogram(mapping = aes(x = recipient_cfscore))

# Beto2014 
ggplot(data = subset(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")) + 
  geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + theme_minimal() + labs(list(title='', x='', y='')) + theme(axis.ticks = element_blank(), axis.text.x = element_blank()) 

# Beto2014 trend line
ggplot(data = subset(contribDB_2014, recipient.name == "O ROURKE, ROBERT (BETO)")) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

# Mitch McC 2014 trend line < -- **TEST FOR THE COMPARISON PLOTS**

ggplot(data = subset(contribDB_2014, recipient.name == "BRAT, DAVID ALAN")) + ggtitle("David Brat 2014/Campaign Funding Strategy") + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 


# grep('SCALISE', contribDB_2014$recipient.name) 
contribDB_2014$recipient.name[660490]

"SCALISE, STEPHEN J"
ggplot(data = subset(contribDB_2014, recipient.name == "SCALISE, STEPHEN J")) + 
  geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + theme_minimal() + labs(list(title='', x='', y='')) + theme(axis.ticks = element_blank(), axis.text.x = element_blank()) 

ggplot(data = subset(contribDB_2014, recipient.name == "SCALISE, STEPHEN J")) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

ggplot(data = subset(contribDB_2014, recipient.name == "PELOSI, NANCY")) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

# Facet Mapping of All Candidates

# FacetTest <- ggplot(contribDB_2014) + geom_point(mapping = aes(x = date, y = contributor.cfscore)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

# Facet Mapping by State
# FacetTest + facet_wrap( ~ recipient.name, nrow=300) + labs(title="Candidate Funding Strategy", caption = "Source: mpg", subtitle="Exploring Differences by Candidates")

# Facet Maping by State

# FacetTest + facet_grid( ~ recipient.state) + labs(title="Candidate Funding Strategy", subtitle="Exploring Funding Strategy Differences by State")

# Comparing Candidates - Cantor v Brant Substitute in Candiate Name as recipient.name to do the comparison 

Can1 <- ggplot(data = subset(contribDB_2014, recipient.name == "CANTOR, ERIC IVAN")) + ggtitle("Eric Cantor (R)\n2014 Campaign Funding Spread") + geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal()  
Can2 <- ggplot(data = subset(contribDB_2014, recipient.name == "BRAT, DAVID ALAN")) + ggtitle("David Brat (D)\n2014 Campaign Funding Spread") + geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

gCan1 <- ggplotGrob(Can1)
gCan2 <- ggplotGrob(Can2)
g <- rbind(gCan1, gCan2, size = "first")
g$widths <- unit.pmax(gCan1$widths, gCan2$widths)
grid.newpage()
grid.draw(g)

# Comparing Candidates - Barber v Scully 2014 

Can1 <- ggplot(data = subset(contribDB_2014, recipient.name == "BARBER, RONALD")) + ggtitle("Ronald Barber (D)\n2014 Campaign Funding Spread") + geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal()  
Can2 <- ggplot(data = subset(contribDB_2014, recipient.name == "MCSALLY, MARTHA")) + ggtitle("Martha McSally (R)\n2014 Campaign Funding Spread") + geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

gCan1 <- ggplotGrob(Can1)
gCan2 <- ggplotGrob(Can2)
g <- rbind(gCan1, gCan2, size = "first")
g$widths <- unit.pmax(gCan1$widths, gCan2$widths)
grid.newpage()
grid.draw(g)

# Comparing Candidates - Barber v Scully 2014 

Can1 <- ggplot(data = subset(contribDB_2014, recipient.name == "BARBER, RONALD")) + ggtitle("Ronald Barber\n2014 Campaign Funding Spread") + geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal()  
Can2 <- ggplot(data = subset(contribDB_2014, recipient.name == "MCSALLY, MARTHA E MS.")) + ggtitle("Martha McSally\n2014 Campaign Funding Spread") + geom_point(mapping = aes(x = date, y = contributor.cfscore, size = amount)) + geom_smooth(mapping = aes(x = date, y = contributor.cfscore), method="lm") + theme_minimal() 

gCan1 <- ggplotGrob(Can1)
gCan2 <- ggplotGrob(Can2)
g <- rbind(gCan1, gCan2, size = "first")
g$widths <- unit.pmax(gCan1$widths, gCan2$widths)
grid.newpage()
grid.draw(g)

```

### Generic Ballot History 
```{r generic ballot}
generic_ballot_history <- read_csv("Political_Data/generic-ballot-history.csv")
View(generic_ballot_history)

ggplot(data = generic_ballot_history) + 
  geom_smooth(mapping = aes(x = end_date, y = Democrat), color="blue") + theme_minimal() + labs(list(title='', x='', y='')) + theme(axis.ticks = element_blank(), axis.text.x = element_blank())

ggplot(data = generic_ballot_history) + theme_minimal() + 
  geom_point(mapping = aes(x = end_date, y = Democrat), color="blue") +   
  geom_smooth(mapping = aes(x = end_date, y = Democrat), color="blue") +   
  geom_point(mapping = aes(x = end_date, y = Republican), color="red") +     
  geom_smooth(mapping = aes(x = end_date, y = Republican), color="red") + 
  geom_point(mapping = aes(x = end_date, y = `Undecided/Other`), color="yellow") + 
  geom_smooth(mapping = aes(x = end_date, y = `Undecided/Other`), color="yellow") + 
  labs(list(title='Generic Polling, 2012-2018', x='', y='% Support'))

ggplot(data = generic_ballot_history) + theme_minimal() + 
  geom_smooth(mapping = aes(x = end_date, y = Democrat), color="blue") +   
  geom_smooth(mapping = aes(x = end_date, y = Republican), color="red") + 
  geom_smooth(mapping = aes(x = end_date, y = `Undecided/Other`), color="yellow") + 
  labs(list(title='Generic Polling, 2012-2018', x='', y='% Support'))
```


